### 2024-09-24 paper/paper.qmd
Updated README and plots in paper

---
title: "When and Where: Examining Crisis Service Calls in Toronto"
subtitle: "A Detailed Look at Temporal and Spatial Trends in Crisis Response Calls"
author: 
  - Xinxiang Gao
thanks: "Code and data are available at: https://github.com/xgao28/PICCFSA."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)
library(gridExtra)
library(cowplot)
```

# Introduction

@citeR, @tidyverse, @ggplot2, @cowplot, @gridExtra

@opendatatoronto

```{r}
#| include: false
#| warning: false
#| message: false
data <- read.csv("../data/analysis_data/piccfsa.csv")
```

# Data {#sec-data}

```{r}
#| label: fig-dow
#| fig-cap: Event Type Counts Grouped by Day of the Week
#| echo: false
#| warning: false
#| message: false
ggplot(data, aes(x = event_dow, fill = event_type)) + 
  geom_bar(position = "dodge") + 
  labs(title = "Event Type Counts Grouped by Day of the Week", 
       x = "Day of the Week", 
       y = "Count",
       fill = "Event Type") +
  theme_minimal()
```

```{r}
#| label: fig-month
#| fig-cap: Event Type Counts Grouped by the Month
#| echo: false
#| warning: false
#| message: false
ggplot(data, aes(x = event_month, fill = event_type)) + 
  geom_bar(position = "dodge") + 
  labs(title = "Event Type Counts Grouped by the Month", 
       x = "Month", 
       y = "Count",
       fill = "Event Type") +
  theme_minimal() +
  coord_flip()
```

```{r}
#| label: fig-year
#| fig-cap: Event Type Counts Grouped by Year
#| echo: false
#| warning: false
#| message: false

ggplot(data, aes(x = event_year, fill = event_type)) + 
  geom_bar(position = "dodge") + 
  labs(title = "Event Type Counts Grouped by Year", 
       x = "Year", 
       y = "Count",
       fill = "Event Type") +
  theme_minimal()
```

```{r}
#| label: fig-hour
#| fig-cap: Event Type Counts Grouped by Hour of the Day
#| echo: false
#| warning: false
#| message: false
df_summarized <- data %>%
  group_by(event_hour, event_type) %>%
  summarise(count = n()) %>%
  ungroup()

# Create the line plot
ggplot(df_summarized, aes(x = event_hour, y = count, color = event_type, group = event_type)) + 
  geom_line() + 
  labs(title = "Event Type Counts Grouped by Hour", 
       x = "Hour", 
       y = "Count", 
       color = "Event Type") +
  theme_minimal()
```

```{r}
#| label: fig-date
#| fig-cap: Event Type Counts Grouped by Date
#| echo: false
#| warning: false
#| message: false
df_summarized <- data %>%
  group_by(event_date, event_type) %>%
  summarise(count = n()) %>%
  ungroup()

# Create the line plot
ggplot(df_summarized, aes(x = event_date, y = count, color = event_type, group = event_type)) + 
  geom_smooth() + 
  labs(title = "Event Type Counts Grouped by Date", 
       x = "Date", 
       y = "Count", 
       color = "Event Type") +
  theme_minimal()
```

```{r}
#| label: fig-hood
#| fig-cap: Event Counts in Neighbourhood Grouped by Event Type
#| echo: false
#| warning: false
#| message: false
# Step 1: Summarize the data by counting occurrences of each event_type
df_summarized <- data %>%
  group_by(hood_158, event_type) %>%
  summarise(count = n()) %>%
  ungroup()

# Step 2: Split the summarized data into separate data frames for each event_type
event_types <- unique(df_summarized$event_type)

# Create a list to hold the top 10 data frames and plots
df_list <- list()
plot_list <- list()
hood_list <- c()

for (event in event_types) {
  # Filter the top 10 neighborhoods for the current event type
  df_top10 <- df_summarized %>%
    filter(event_type == event) %>%
    arrange(desc(count)) %>%
    head(10)

  # Save the data frame to the list
  df_list[[event]] <- df_top10
  hood_list <- c(hood_list, df_top10$hood_158)

  # Step 3: Create a bar plot for the top 10 neighborhoods for the current event type
  plot <- ggplot(df_top10, aes(x = reorder(hood_158, count), y = count)) + 
    geom_bar(stat = "identity") + 
    labs(title = event, 
         x = "Neighborhood 158", 
         y = "Count") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    coord_flip()

  # Save the plot to the list
  plot_list[[event]] <- plot
}

# Step 4: Combine the individual plots using cowplot
combined_plot <- cowplot::plot_grid(plotlist = plot_list, ncol = 3)

# Step 5: Display the combined plot
print(combined_plot)
```

```{r}
#| label: tab_hood
#| fig-cap: A Look-up Table for Hood Number and Hood Name
#| echo: false
#| warning: false
#| message: false
# Step: Get a look-up table
hood_set <- unique(hood_list)
neighborhood_table <- data %>%
  select(hood_158, neighbourhood_158) %>%
  distinct() %>%
  filter(hood_158 %in% hood_set) %>%
  arrange(hood_158) %>%
  rename("hood number" = hood_158,  
         "hood name" = neighbourhood_158)

grid.arrange(grid.table(neighborhood_table))
```

\newpage

# References