### 2024-09-27 paper/paper.qmd
Added removing NA observations
and updated accordingly

---
title: "When and Where: Understanding the Patterns of Emergency Service Calls in Toronto"
subtitle: "Exploring the Temporal and Spatial Dynamics of Reported Emergency Incidents"
author: 
  - Xinxiang Gao
thanks: "Code and data are available at: https://github.com/xgao28/PICCFSA."
date: today
date-format: long
toc: true
abstract: "This paper examines ten years of data on crisis service calls in Toronto, focusing on overdose, Person in Crisis, and suicide-related incidents. Our analysis reveals clear temporal patterns, with overdoses peaking at night and mental health crises spiking in the afternoon. Spatially, certain neighborhoods such as Moss Park and Downtown Yonge East consistently report higher numbers of crisis events. These findings highlight the importance of targeted mental health and substance abuse interventions, as well as resource allocation strategies tailored to specific times and locations."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)
library(gridExtra)
library(cowplot)
library(knitr)
```

# Introduction

The examination of crisis service calls is critical for enhancing public safety and effectively supporting vulnerable populations in urban environments. As mental health issues and substance use continue to rise, it is essential for emergency response services to understand the nature and patterns of crisis situations. Analyzing when and where these calls occur can inform strategic planning, resource allocation, and intervention methods for law enforcement and mental health services. By shedding light on the dynamics of crisis response, we can better equip first responders to address the complex needs of individuals in distress, ultimately fostering safer and more resilient communities.

This study aims to analyze the temporal and spatial trends of crisis-related incidents in Toronto, as recorded in the Persons in Crisis Calls for Service dataset. The primary objectives are to:

- Identify the most common types of crisis events and their distribution across neighborhoods.
- Examine the temporal patterns of these events, including daily, weekly, and seasonal trends.

The sections are organized as follows: 
@sec-data provides a detailed description of the data and its sources, including the context and ethical considerations. @sec-results presents the results using visualizations as well as discusses the implications of the findings, addressing both the practical applications and the limitations of the study. Finally, @sec-conclusion concludes the paper and suggests directions for future research.

# Data {#sec-data}

The Persons in Crisis Calls for Service dataset, provided by the Toronto Police Service (TPS), offers a comprehensive view of crisis-related incidents attended by TPS. This dataset includes over 309,000 records from January 1, 2014, to June 30, 2024, focusing on significant event types such as Suicide, Person in Crisis, and Overdose. The dataset captures incidents that required police intervention, excluding those handled by other TPS divisions, such as Parking and Marine. Additionally, the dataset does not contain sensitive information related to individuals in crisis, ensuring that no individual can be identified from the results.

Understanding the nature and patterns of crisis calls is foundational for several reasons. For public safety, identifying high-activity areas and time periods can help in deploying resources more effectively, ensuring quicker response times and better support for individuals in crisis.In terms of resource allocation, statistical results towards the demand for crisis services can inform budgeting and staffing decisions, ensuring that resources are allocated where they are most needed.For policy development, the findings can guide the development of policies and interventions aimed at reducing the incidence of crises and improving community well-being.

To provide a comprehensive overview of the data, we use a combination of text, tables, and visualizations. The following sections describe the key characteristics of the dataset and summarize the main findings.

# Results {#sec-results}

```{r}
#| include: false
#| warning: false
#| message: false
data <- read.csv("../data/analysis_data/piccfsa.csv")

# Factor day of week and month for correct ordering
data$event_dow <- factor(data$event_dow, 
                         levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))

month_levels <- c("January", "February", "March", "April", "May", "June",
                  "July", "August", "September", "October", "November", "December")

data$event_month <- factor(data$event_month, levels = month_levels)
```

```{r summary_stats}
#| label: tbl-summary
#| echo: false
#| warning: false
#| message: false
summary_stats <- data %>%
  group_by(event_type) %>%
  summarise(
    total_count = sum(n())
  ) %>%
  ungroup() %>% 
  rename("Event Type" = event_type, "Total Count" = total_count)

kable(summary_stats, caption = "Summary Statistics of Crisis Events", align = 'l', booktabs = TRUE)
```

## Overview
Over the ten-year period, the dataset reveals the nature and frequency of three primary types of crisis events: Overdose, Person in Crisis, and Suicide-related, as summarized in @tbl-summary. Among these, Person in Crisis events are by far the most frequent, comprising more than half of the total incidents.

- Overdose: There were a total of 32,494 overdose incidents recorded over the decade. This category includes all cases where an individual required police intervention due to drug or alcohol overdose. The number of overdose incidents highlights the ongoing public health challenge related to substance abuse in the community.

- Person in Crisis: With 172,149 incidents, Person in Crisis is the most frequently reported type of event. These incidents involve individuals who are experiencing mental health crises or other forms of distress that necessitate police involvement. The high count underscores the critical need for mental health support and resources within the community.

- Suicide-related: A total of 92,762 suicide-related incidents were documented. This includes both attempts and situations where there was a risk of suicide, requiring immediate police response. The prevalence of such incidents emphasizes the importance of suicide prevention efforts and access to mental health services.

## Temporal Analysis

```{r}
#| label: fig-hour
#| fig-cap: Event Type Counts by Hour of the Day
#| echo: false
#| warning: false
#| message: false
df_summarized <- data %>%
  group_by(event_hour, event_type) %>%
  summarise(count = n()) %>%
  ungroup()

# Create the line plot
ggplot(df_summarized, aes(x = event_hour, y = count, color = event_type, group = event_type)) + 
  geom_line() + 
  labs(title = "Event Type Counts by Hour", 
       x = "Hour", 
       y = "Count", 
       color = "Event Type") +
  theme_minimal()
```

@fig-hour shows the distribution of event types across the hours of the day. This plot helps identify any daily patterns in the occurrence of crisis events. Overdose incidents, represented in red, peak at night, highlighting the potential need for heightened availability of medical and police services during these hours. Person in Crisis events, shown in green, are most frequent in the afternoon and steadily decline toward evening, suggesting that mental health resources may need to be more accessible during business hours. Suicide-related events, depicted in blue, follow a trend similar to Overdose events but with lower overall counts, peaking in the late afternoon and early evening hours.

```{r}
#| label: fig-dow
#| fig-cap: Event Type Counts by Day of the Week
#| echo: false
#| warning: false
#| message: false

ggplot(data, aes(x = event_dow, fill = event_type)) + 
  geom_bar(position = "dodge") + 
  labs(title = "Event Type Counts by Day of the Week", 
       x = "Day of the Week", 
       y = "Count",
       fill = "Event Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

@fig-dow illustrates the distribution of event types across the days of the week, showing that Person in Crisis events (green bars) consistently dominate the count across all days. Thursdays have the highest count of all events, while weekends see a marked drop. This could reflect workplace or social stress accumulating during the week, leading to crises toward the week's end. The reduction during weekends may point to fewer reports or different response patterns, offering an opportunity for further investigation.

```{r}
#| label: fig-month
#| fig-cap: Event Type Counts by the Month
#| echo: false
#| warning: false
#| message: false
ggplot(data, aes(x = event_month, fill = event_type)) + 
  geom_bar(position = "dodge") + 
  labs(title = "Event Type Counts by the Month", 
       x = "Month", 
       y = "Count",
       fill = "Event Type") +
  theme_minimal() +
  coord_flip()
```

@fig-month displays the distribution of event types across the months of the year, revealing seasonal trends. Person in Crisis events (green bars) consistently have the highest counts throughout the year, followed by Suicide-related events (blue bars). Overdose events (red bars) are the least frequent. May and June see the highest number of events across all categories, while February has the lowest, potentially due to fewer outdoor activities during colder months. These trends may reflect how seasonal factors and societal conditions influence crisis situations.

```{r}
#| label: fig-date
#| fig-cap: Event Type Counts Over Time
#| echo: false
#| warning: false
#| message: false
df_summarized <- data %>%
  group_by(event_date, event_type) %>%
  summarise(count = n()) %>%
  ungroup()

df_summarized$event_date <- as.Date(df_summarized$event_date, format = "%Y-%m-%d")

ggplot(df_summarized, aes(x = event_date, y = count, color = event_type, group = event_type)) + 
  geom_smooth() + 
  labs(title = "Event Type Counts Over Time", 
       x = "Date", 
       y = "Count", 
       color = "Event Type") +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y",
               limits = as.Date(c(min(df_summarized$event_date), max(df_summarized$event_date)))) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

@fig-date presents the temporal trends of crisis calls over the entire period. Overall, the chart indicates a general upward trend in the number of crisis events, with a significant peak during the pandemic years of 2021 and 2022, coinciding with lockdowns and restrictions. This surge likely reflects the psychological toll of isolation and economic stress, contributing to more frequent mental health crises and overdoses. Another noticeable increase is seen in 2024. These patterns underscore the importance of mental health and substance abuse support during times of social upheaval.

## Spatial analysis

```{r}
#| label: fig-hood
#| fig-cap: Event Counts in Neighbourhood by Event Type
#| echo: false
#| warning: false
#| message: false
# Step 1: Summarize the data by counting occurrences of each event_type
df_summarized <- data %>%
  group_by(hood_158, event_type) %>%
  summarise(count = n()) %>%
  ungroup()

# Step 2: Split the summarized data into separate data frames for each event_type
event_types <- unique(df_summarized$event_type)

# Create a list to hold the top 10 data frames and plots
df_list <- list()
plot_list <- list()
hood_list <- c()

# Define colors for each event type
event_colors <- c("Overdose" = "#1B9E77",      # Deep green
                  "Person in Crisis" = "#D95F02",  # Muted orange
                  "Suicide-related" = "#7570B3")   # Slate blue

for (event in event_types) {
  # Filter the top 10 neighborhoods for the current event type
  df_top10 <- df_summarized %>%
    filter(event_type == event) %>%
    arrange(desc(count)) %>%
    head(10)

  # Save the data frame to the list
  df_list[[event]] <- df_top10
  hood_list <- c(hood_list, df_top10$hood_158)

  # Step 3: Create a bar plot with a consistent color for each event type
  plot <- ggplot(df_top10, aes(x = reorder(hood_158, count), y = count)) + 
    geom_bar(stat = "identity", fill = event_colors[event]) +  # Set a fixed color for the bars
    labs(title = event, 
         x = "Neighborhood 158", 
         y = "Count") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    coord_flip()

  # Save the plot to the list
  plot_list[[event]] <- plot
}

# Step 4: Combine the individual plots using cowplot
combined_plot <- cowplot::plot_grid(plotlist = plot_list, ncol = 3)

# Step 5: Display the combined plot
print(combined_plot)
```

```{r tab_hood}
#| label: tbl-hood
#| echo: false
#| warning: false
#| message: false
# Step: Get a look-up table
hood_set <- unique(hood_list)
neighborhood_table <- data %>%
  select(hood_158, neighbourhood_158) %>%
  distinct() %>%
  filter(hood_158 %in% hood_set) %>%
  arrange(hood_158) %>%
  rename("hood number" = hood_158,  
         "hood name" = neighbourhood_158)

kable(neighborhood_table, caption = "Hood Number and Name Reference", 
             booktabs = TRUE, align = 'l', label = "tab-hood")
```

@fig-hood showcases the top 10 neighborhoods with the highest counts for each event type. Moreover, @tbl-hood provides a lookup table for the neighborhood numbers and their corresponding names, which can be used to interpret the neighborhood codes in the previous plots. To illustrate, Moss Park and Downtown Yonge East have the most reported cases for all three event types, with Overdose incidents dominating other neighborhoods. The high counts in these areas may be linked to population density, socioeconomic challenges, or proximity to services like shelters and social support networks. Kensington-Chinatown ranks third for Person in Crisis events, while Annex ranks third for Suicide-related events. These findings suggest that certain neighborhoods may require targeted interventions based on the specific crisis patterns observed.

# Conclusion {#sec-conclusion}

The analysis of the Persons in Crisis Calls for Service dataset shows notable patterns in the timing and location of crisis-related incidents in Toronto over the past decade. Overdose, Person in Crisis, and Suicide-related events each follow distinct daily, weekly, and seasonal cycles. Overdoses tend to occur most frequently at night, Person in Crisis calls are most common in the afternoon, and Suicide-related incidents maintain a steady rate during late afternoon and early evening. Thursdays experience the highest volume of crisis calls across all types, while weekends typically have fewer incidents. These time-based trends help guide the allocation of resources and the adjustment of response plans to align with periods of greatest demand.

In terms of location, areas like Moss Park and Downtown Yonge East consistently report high numbers of crisis calls, particularly for overdoses. These findings suggest the need for focused interventions in these neighborhoods, potentially influenced by socioeconomic conditions or population density. Recognizing both the temporal and geographic patterns of crisis incidents is key to enhancing emergency response strategies and supporting the areas most impacted.

Knowing when and where these incidents are most likely to happen is vital for planning emergency response efforts and distributing resources effectively. By identifying neighborhoods and time periods with higher incident rates, this analysis offers practical guidance for policymakers and service providers to improve crisis management, ultimately contributing to public safety and community well-being.

The dataset also includes additional information that could deepen the analysis, such as police division details, whether an official occurrence was logged, and whether a Mental Health Act (MHA) apprehension occurred. Future studies could examine how these elements influence crisis resolution. Moreover, although this dataset focuses on three crisis types, incorporating a broader range of emergency call categories could provide a more detailed understanding of Toronto's crisis response landscape.

# Acknowledgements

We would like to express our gratitude to the developers and contributors of R [@citeR] as well as several R packages that were essential for the analysis and visualization of the data in this report. The following R packages provided indispensable tools and functionalities:

- `tidyverse` [@tidyverse]: A collection of R packages designed for data science, including dplyr, ggplot2, readr, purrr, and others, which greatly facilitated data manipulation, analysis, and visualization.

- `ggplot2` [@ggplot2]: An implementation of the Grammar of Graphics, which allowed us to create complex and aesthetically pleasing visualizations with ease.

- `cowplot` [@cowplot]: The package enabled us to combine multiple plots into a single figure, enhancing the presentation of our results.

- `lubridate` [@lubridate]: This package facilitates the work with date and time data, simplifying the process of manipulating and analyzing temporal information.

- `knitr` [@knitr]: This package enabled us to perform data demonstration with tables.

- `janitor` [@janitor]: This package provided helpful tools for data cleaning. Its functionalities greatly enhanced the efficiency and accuracy of our data preparation process.

Additionally, we are grateful to the City of Toronto and the Open Data Portal for providing access to the dataset through the opendatatoronto package [@opendatatoronto]. The availability of such high-quality, open-source data is vital for conducting meaningful and impactful research.

Finally, we would like to thank all those who contributed to the development and maintenance of the R programming language and its ecosystem, as well as the broader open-source community, whose efforts make such research possible.

# Appendix

## Data Cleaning

In the data cleaning process, the raw dataset was first read from a CSV file and unnecessary columns, including identifiers and irrelevant variables, were removed. The janitor package was used to clean the column names for consistency. Instances of the value "NSA" in character columns were replaced with NA, and any rows containing NA values were subsequently dropped. To ensure proper ordering in visualizations, the day of the week and month were converted to factors with defined levels. Finally, the cleaned dataset was saved as a new CSV file for further analysis. The cleaned dataset contains over 297,000 observations of 10 variables.

\newpage

# References